---
title: Reproducible Research 
output:
  html_document:
  md_document:
    variant: markdown_github
---

Week 2 Course Project 1
=========================
install and library packages
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

1. Load data activity.csv
```{r}
activity<-read.csv("activity.csv")
activity_nna<-activity %>% 
  na.omit()
str(activity)
summary(activity)
activity_byday<-activity_nna %>% 
    group_by(date) %>% 
    summarise(sum_steps=sum(steps))

```

2. Histogram of the total number of steps taken each day
```{r}
ggplot(data=activity_byday,aes(x=sum_steps)) +
  geom_histogram()+
  xlab("Total number of steps each day")
```
  
3. Mean and median number of steps taken each day
```{r}
mean<-mean(activity_byday$sum_steps)
median<-median(activity_byday$sum_steps)
paste("Mean number of steps each day is", mean, sep=" ")
paste("Median number of steps each day is", median, sep=" ")
```
4. Time series plot of the average number of steps taken
```{r}
activity_interval<-activity_nna %>% 
    group_by(interval) %>% 
    summarise(mean_steps=mean(steps))
ggplot(data=activity_interval,aes(x=interval, y=mean_steps)) +
  geom_line()
```

5. The 5_minute interval that contains the max number of steps
```{r}
max_steps<-max(activity_interval$mean_steps)
max_interval<-activity_interval[activity_interval$mean_steps==max_steps,]$interval
paste('Max steps is', max_steps, sep=" ")
paste('Max interval is',max_interval, sep=" ")
```

6. Imputing missing data
Since there are missing data for some of the days, it is not feasible to impute the missing values with the mean steps that day. So I use the mean steps for the 5-minute interval. The data is available in activity_merge dataset.
```{r}
#number of rows with missing values
nrow(activity)-nrow(activity_nna)
#merge the dataset with mean steps for the 5 minute interval with the original dataset
activity_merge<-merge(activity,activity_interval)
#Relace NA with the mean steps for the 5 minute interval
activity_merge$steps[is.na(activity_merge$steps)]<-activity_merge$mean_steps[is.na(activity_merge$steps)]
```

7. Histogram of total number os steps taken after imputation of missing values
```{r}
activity_byday2<-activity_merge %>% 
    group_by(date) %>% 
    summarise(sum_steps=sum(steps))
ggplot(data=activity_byday2,aes(x=sum_steps)) +
  geom_histogram()+
  xlab("Total number of steps each day")+
  ggtitle("Total number of steps each day after missing values are imputated")
#mean and median total number of steps per day
mean(activity_byday2$sum_steps)
median(activity_byday2$sum_steps)
# mean and median total number of steps per day are different from that calculated before missing values have been imputed
```

8. Panel plot comparing the average number of stps taken per 5-minute interval across weekdays and weekends
```{r}
#add new variable "day" with two levels-"Weekday" and "Weekend"
activity_merge$weekday<-weekdays(as.Date(activity_merge$date))
activity_merge$day<-ifelse(activity_merge$weekday=="Saturday"|activity_merge$weekday=="Sunday","Weekend","Weekday")
#calculate average steps by interval and weekday/weekend
activity_new<-activity_merge %>% 
    group_by(day,interval) %>% 
    summarise(mean_steps=mean(steps))
#make the plot
ggplot(data=activity_new,aes(x=interval, y=mean_steps,color=day)) +
  geom_line() +
  facet_grid(day ~ .)+
  labs(title="Average Steps by Interval and Weekday/Weekend", x="5-Minute Interval", y="Average steps")
```

